---
title: "Faketucky Analysis"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    source: embed
---

```{r setup, include=FALSE}
library(flexdashboard)
library(readr)
library(tidyverse)
library(stringr)

# clean data ####
ft_data <- read_csv("data/faketucky_20160921.csv")
```


### What do high school GPA's look like in Faketucky?

```{r}

ggplot(ft_data, aes(x = avg_gpa_hs))+
  geom_histogram()+
  theme_bw()+
  theme(legend.position = "none",
        text = element_text(size = 10))
```
 
***

The distribtion is mostly between 2.0 and 3.0, so there doesn't appear to be a significant amount of grade inflation.

### Does the distribution of GPA's differ by urbanicity?

```{r}

ggplot(ft_data, aes(x = avg_gpa_hs))+
  geom_histogram(binwidth = .3)+
  facet_wrap(~first_hs_urbanicity)+
  theme_bw()+
  theme(legend.position = "none",
        text = element_text(size = 10))
```

***

This chart makes it hard to tell what's going on in urbanicities with a smaller n of students, but we can see the distribution of the most common urbanicities fairly well.


### Can we get a better look at urbanicities with a smaller n count?

```{r}
ggplot(ft_data, aes(x = avg_gpa_hs))+
  geom_histogram(binwidth = .3)+
  facet_wrap(~first_hs_urbanicity, scales = "free_y")+
  theme_bw()+
  theme(legend.position = "none",
        text = element_text(size = 10))
```

***

Allowing for a "floating y axis" allows us to see the distrubtion in each urbanicity more clearly.

It looks like there may be some grade inflation in:

- `City: Small`
- `Town: Fringe`

### Is there a correlation between HS GPA and Math ACT score?

```{r}
ggplot(ft_data, aes(x = avg_gpa_hs, scale_score_11_math))+
  geom_point(alpha = .2)+
  geom_smooth(method = "lm")+
  geom_hline(yintercept = 19, color = "grey40")+
  facet_wrap(~first_hs_urbanicity)+
  theme_bw()+
  theme(legend.position = "none",
        text = element_text(size = 10))
```

***

Yep - there's a relationship! 

It looks pretty linear, but there's a lot of noise. I wonder if there's a way to get a better idea of what's going on here...

### Density plots & a GAM can eliminate noisy large-n data
```{r}
ggplot(ft_data, aes(x = avg_gpa_hs, scale_score_11_math))+
  stat_density2d(aes(fill = ..level..,
                     alpha = ..level..),
                 size = 2, bins = 4,
                 geom = "polygon")+
  geom_smooth(se = FALSE)+
  geom_hline(yintercept = 19, color = "goldenrod4")+
  theme_bw()+
  theme(legend.position = "none",
        text = element_text(size = 10))

```

***

Woah! This paints a different picture than using `geom_point()` to plot the data. Turns out getting a really good GPA is a pretty darn good indication you're going to meet the KYCPE math benchmark!


 
### Density of GPA & Math ACT score with a GAM regression

```{r}
ggplot(ft_data, aes(x = avg_gpa_hs, scale_score_11_math))+
  stat_density2d(aes(fill = ..level..,
                     alpha = ..level..),
                 size = 2, bins = 4,
                 geom = "polygon")+
  geom_smooth(se = FALSE)+
  geom_hline(yintercept = 19, color = "goldenrod4")+
  facet_wrap(~first_hs_urbanicity)+
  theme_bw()+
  theme(legend.position = "none",
        text = element_text(size = 10))

```

***

Maybe there isn't grade inflation in `City: Small` or `Town: Fringe`. They could have:

- A high concentration of high-performing students
- Good programs to support gifted students

It would be interesting to investigate districts in these settings more closely to get a better idea of what's happening in these areas.